---
title: "p8105_hw3_ml4198"
author: "Matthew Lawlor"
date: "10/12/2020"
output: github_document
---

```{r setup, include=FALSE}
library (tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
  
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

```{r data}
library(p8105.datasets)
data("instacart")
```

This is a dataframe of Instacart orders with 17 variables and 1,384,617 entries. Key variables are user_id, order_id, product_id, aisle_id.

```{r}
instacart %>% 
    group_by(aisle_id) %>% 
  summarize(
    n_obs = n()
    ) %>% 
  mutate(
    freq_rank = min_rank(desc(n_obs))
    ) %>% 
  filter(freq_rank < 10)
```

There are 134 aisles, and the most commonly ordered from are aisle 83 (150609 orders), 24 (150473 orders), and 123 (78493 orders).

**Code for figure 1**

```{r}
instacart %>% 
    group_by(aisle_id) %>% 
  summarize(
    n_obs = n()
    ) %>% 
  mutate(
    freq_rank = min_rank(desc(n_obs))
    ) %>% 
  filter(n_obs > 10000) %>% 
  arrange(freq_rank) %>% 
  ggplot(aes(x = freq_rank, y = n_obs, color = aisle_id)) +
  geom_point() + 
  scale_x_continuous(
    breaks = c(1, 5, 10, 15, 20, 25, 30, 35),
    minor_breaks = c(1:39)) +
  labs(
    x = "Aisle Rank",
    y = "Number of items ordered",
    title = "Figure 1",
    color = "Aisle ID"
  )
```

**Code for table 1**

```{r}
instacart %>% 
  select(aisle, product_name) %>% 
  filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_obs = n()) %>% 
  arrange(aisle, desc(n_obs)) %>% 
  filter(min_rank(desc(n_obs)) < 4)
```
  
**Code for Table 2**

```{r}
instacart %>% 
  select(order_dow, order_hour_of_day, product_name) %>% 
  filter(product_name == c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = product_name,
    values_from = mean_hour
  ) %>% 
  knitr::kable(digits = 1)
```

## Problem 2

```{r}
accel_df = read.csv("./accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440, 
    names_to = "minute_of_day",
    names_prefix = "activity_",
    values_to = "activity_counts") %>% 
  mutate(
    minute_of_day = as.integer(minute_of_day),
    weekend = ifelse(grepl("Saturday", day), TRUE,
    ifelse(grepl("Sunday", day), TRUE, FALSE))) %>% 
  select(week, day_id, day, weekend, everything())
```

accel_df is a dataframe with 6 variables and 50,400 entries. Variables include week and day of study period, day of week, weekend, minute of day, and activity counts per minute. 

**Table**

```{r}
accel_df %>% 
  group_by(day_id, day) %>% 
  summarize(sum(activity_counts)) %>% 
  knitr::kable(digits = 0)
```

**Plot**

```{r}
accel_df %>% 
  ggplot(aes(x = minute_of_day, y = activity_counts, color = day)) +
  geom_point(alpha = .5)
```

There is a trimodal distribution of activity pattern throughout the day, most notably in morning and evening, less pronounced at midday.

##Problem 3

```{r}
library(p8105.datasets)
data("ny_noaa")
```

```{r}
ny_noaa_tidy =
ny_noaa %>% 
  separate(date, into = c("year", "mon", "day")) %>% 
  mutate(
    prcp = prcp / 100,
    snow = snow / 10,
    snwd = snwd /10,
    tmax = as.numeric(tmax),
    tmin = as.numeric(tmin),
    tmax = tmax / 10,
    tmin = tmin / 10,
    mon = as.integer(mon)
  ) 
```

This is a dataframe with variables and 2,595,176 entries. Key variables are id, date, precipitation. Data completeness is high in precipitation, snow, and snow depth (94%, 85%, and 77%, respectively), but relatively poor in tmax and tmin (56% each).

**Plot 1**

```{r}
ny_noaa_tidy %>% 
  filter(mon == 1 | mon == 7) %>%
  group_by(mon, id) %>% 
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) %>% 
  ggplot(aes(x = mean_tmax)) +
  geom_histogram() +
  facet_grid(. ~ mon)
```

Mean tmax across weather stations appear normally distributed around 0 deg C in January and 28 deg C in July.

**Plot 2**

```{r}
ny_noaa_tidy %>%
  ggplot(aes(x = tmax, y = tmin)) +
  geom_bin2d()
```

